\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2022-02-07 2-the-cubble-pkg.Rmd
add some sentences on data cube and fix \code{...} rendering

Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this. No matter which form the data is in, these formats share some common components that characterise the spatio-temporal data. A spatial identifier (\code{id} in the diagram) is the unique identifier of each site. The temporal identifier (\code{t} in the diagram) prescribes the time stamp each site is recorded. Coordinates, comprising of latitude and longitude (\code{lat} and \code{(long)} in the diagram), locates each site on the map. These identifiers will be the building blocks for the data structure introduced below. Other variables in the data can be categorised into two groups: spatial variables that are invariant at each time stamp for every site, i.e. the name or code of the weather station, and temporal variables that varies with time.

@@

`as_cubble()} creates a \pkg{cubble} by supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude first and then latitude. The naming of \code{key} and \code{index} follows the convention in the \pkg{tsibble} package. The cubble created by default is in the nested form. Below is an example of creating a cubble:

@@

There are a few information in the \pkg{cubble} header: the name of the \code{key} variable (`id`) and its number (5), bounding box (`[115.97, -32.94, 133.55, -12.42]`), and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.

@@

Building from an \code{tbl_df} class, \pkg{cubble} has implemented methods for \pkg{dplyr} generics, which includes:

>>>>

Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this. These four examples show some common characters of the spatio-temporal data: 1) a spatial identifier, that uniquely identifies each site, 2) a temporal identifier, that prescribes the time stamp each site is recorded, 3) Coordinates, comprising of latitude and longitude, that locate each site on the map. These identifiers will be the building blocks for cubble. Spatio-temporal data can also be thought of as a data cube with three axes: id, time, and variable. Notice that in the time-variable face, time invariant variables share the same value across all the time stamps while time varying variables have values varying across time. This heterogeneity will be used to split the data into two forms in cubble.

@@

\code{as_cubble()} creates a cubble by supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude first and then latitude. The naming of \code{key} and \code{index} follows the convention in the \pkg{tsibble} package. The cubble created by default is in the nested form. Below is an example of creating a cubble:

@@

There are a few information in the cubble header: the name of the \code{key} variable (`id`) and its number (5), bounding box (`[115.97, -32.94, 133.55, -12.42]`), and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.

@@

Building from an \code{tbl_df} class, cubble has implemented methods for \pkg{dplyr} generics, which includes:

### 2022-02-07 3-advanced.Rmd
add some sentences on data cube and fix \code{...} rendering

`match_sites()} from \pkg{cubble} matches data based on these two criteria. It first matches the two data sources spatially through computing the pairwise distance based on latitude and longitude. Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window. Figure \ref{fig:illu-matching} illustrate this temporal matching in more details. Given two series \code{A} and \code{a}. 3 peaks have been picked in each series. An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals. In this illustration, there are 2 matches for these two series. Several arguments are available in \code{match_sites()} to fine-tune the matching:

@@

Glyph map [@Wickham2012-yr] plots the time series as single glyph on the map and in \proglang{R}, \pkg{GGally} implements the glyph map through the \code{glyphs()} function. The function construct a data frame with calculated position (`gx}. \code{gy}. \code{gid}. of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr). The data can then be piped into \code{ggplot} to create the glyph map as:

@@

Polar glyph map can be specify as a parameter, \code{polar = TRUE}. in the \code{geom_glyph()}. along with \code{width} and \code{height} in either absolute or relative value. Global and local scale can be controlled by the parameter \code{global_rescale}. which default to \code{TRUE} for global scaling. Reference box and line can be added with separate \code{geom_glyph_box()} and \code{geom_glyph_line()`.

>>>>

\code{match_sites()} from \pkg{cubble} matches data based on these two criteria. It first matches the two data sources spatially through computing the pairwise distance based on latitude and longitude. Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window. Figure \ref{fig:illu-matching} illustrate this temporal matching in more details. Given two series \code{A} and \code{a}. 3 peaks have been picked in each series. An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals. In this illustration, there are 2 matches for these two series. Several arguments are available in \code{match_sites()} to fine-tune the matching:

@@

Glyph map [@Wickham2012-yr] plots the time series as single glyph on the map and in \proglang{R}, \pkg{GGally} implements the glyph map through the \code{glyphs()} function. The function construct a data frame with calculated position (\code{gx}. \code{gy}. \code{gid}. of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr). The data can then be piped into \code{ggplot} to create the glyph map as:

@@

Polar glyph map can be specify as a parameter, \code{polar = TRUE}. in the \code{geom_glyph()}. along with \code{width} and \code{height} in either absolute or relative value. Global and local scale can be controlled by the parameter \code{global_rescale}. which default to \code{TRUE} for global scaling. Reference box and line can be added with separate \code{geom_glyph_box()} and \code{geom_glyph_line()}.

### 2022-02-07 4-4-netcdf.Rmd
add some sentences on data cube and fix \code{...} rendering

```{r eval = FALSE, echo = TRUE}
# Assume \code{my_ncdf} has bounding box of [-180, -90, 180, -90] 
# at 0.25 degree resolution and subset it to have 
# 1 degree resolution:
dt <- as_cubble(my_ncdf, vars = c("q", "z"), 
                long_range = seq(-180, 180, 1), 
                lat_range = seq(-90, 90, 1))
```

>>>>

```{r eval = FALSE, echo = TRUE}
# Assume my_ncdf has bounding box of [-180, -90, 180, -90] 
# at 0.25 degree resolution and subset it to have 
# 1 degree resolution:
dt <- as_cubble(my_ncdf, vars = c("q", "z"), 
                long_range = seq(-180, 180, 1), 
                lat_range = seq(-90, 90, 1))
```

### 2022-02-07 main.tex
add some sentences on data cube and fix \code{...} rendering

Spatio-temporal data usually come in various forms and Figure
\ref{fig:illu-input} shows four examples of this. No matter which form
the data is in, these formats share some common components that
characterise the spatio-temporal data. A spatial identifier (\code{id}
in the diagram) is the unique identifier of each site. The temporal
identifier (\code{t} in the diagram) prescribes the time stamp each site
is recorded. Coordinates, comprising of latitude and longitude
(\code{lat} and \code{(long)} in the diagram), locates each site on the
map. These identifiers will be the building blocks for the data
structure introduced below. Other variables in the data can be
categorised into two groups: spatial variables that are invariant at
each time stamp for every site, i.e.~the name or code of the weather
station, and temporal variables that varies with time.

@@

`as\_cubble()\} creates a \pkg{cubble} by supplying the three key
components: \code{key} as the spatial identifier; \code{index} as the
temporal identifier; and a vector of \code{coords} in the order of
longitude first and then latitude. The naming of \code{key} and
\code{index} follows the convention in the \pkg{tsibble} package. The
cubble created by default is in the nested form. Below is an example of
creating a cubble:

@@

There are a few information in the \pkg{cubble} header: the name of the
\code{key} variable (\texttt{id}) and its number (5), bounding box
(\texttt{{[}115.97,\ -32.94,\ 133.55,\ -12.42{]}}), and also the name of
variable nested in the \code{ts} column with its type:
\code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.

@@

`match\_sites()\} from \pkg{cubble} matches data based on these two
criteria. It first matches the two data sources spatially through
computing the pairwise distance based on latitude and longitude. Pairs
that pass the spatial matching are then matched temporally through
computing the number of matched peaks within a fixed length window.
Figure \ref{fig:illu-matching} illustrate this temporal matching in more
details. Given two series \code{A} and \code{a}. 3 peaks have been
picked in each series. An interval, with default length of 5, is
constructed for each peak in series \code{A} and the peaks in series
\code{a} are tested against whether they fall into the any of the
intervals. In this illustration, there are 2 matches for these two
series. Several arguments are available in \code{match_sites()} to
fine-tune the matching:

@@

Glyph map \citep{Wickham2012-yr} plots the time series as single glyph
on the map and in \proglang{R}, \pkg{GGally} implements the glyph map
through the \code{glyphs()} function. The function construct a data
frame with calculated position (`gx\}. \code{gy}. \code{gid}. of each
point on the time series using linear algebra (Equation 1 and 2 in
\citet{Wickham2012-yr}). The data can then be piped into \code{ggplot}
to create the glyph map as:

@@

Polar glyph map can be specify as a parameter, \code{polar = TRUE}. in
the \code{geom_glyph()}. along with \code{width} and \code{height} in
either absolute or relative value. Global and local scale can be
controlled by the parameter \code{global_rescale}. which default to
\code{TRUE} for global scaling. Reference box and line can be added with
separate \code{geom_glyph_box()} and
\textbackslash code\{geom\_glyph\_line()`.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # Assume \code{my_ncdf} has bounding box of [-180, -90, 180, -90] 
R> # at 0.25 degree resolution and subset it to have 
R> # 1 degree resolution:
R> dt <- as_cubble(my_ncdf, vars = c("q", "z"), 
+                 long_range = seq(-180, 180, 1), 
+                 lat_range = seq(-90, 90, 1))
\end{CodeInput}
\end{CodeChunk}

@@

With spatio-temporal data, users may wish to make plots to learn the
spatial distribution of a variable or to find patterns such as trend or
seasonality in the time series. Combining this two types of plot with
interactivity let users to link between points on the map and their
corresponding time series and explore the spatial and temporal dimension
of the data simultaneously. Below is an example that describes the
process of building an interactive graphic with \pkg{cubble} and
\textbackslash pkg\{crosstalk\{.

>>>>

Spatio-temporal data usually come in various forms and Figure
\ref{fig:illu-input} shows four examples of this. These four examples
show some common characters of the spatio-temporal data: 1) a spatial
identifier, that uniquely identifies each site, 2) a temporal
identifier, that prescribes the time stamp each site is recorded, 3)
Coordinates, comprising of latitude and longitude, that locate each site
on the map. These identifiers will be the building blocks for cubble.
Spatio-temporal data can also be thought of as a data cube with three
axes: id, time, and variable. Notice that in the time-variable face,
time invariant variables share the same value across all the time stamps
while time varying variables have values varying across time. This
heterogeneity will be used to split the data into two forms in cubble.

@@

\code{as_cubble()} creates a cubble by supplying the three key
components: \code{key} as the spatial identifier; \code{index} as the
temporal identifier; and a vector of \code{coords} in the order of
longitude first and then latitude. The naming of \code{key} and
\code{index} follows the convention in the \pkg{tsibble} package. The
cubble created by default is in the nested form. Below is an example of
creating a cubble:

@@

There are a few information in the cubble header: the name of the
\code{key} variable (\texttt{id}) and its number (5), bounding box
(\texttt{{[}115.97,\ -32.94,\ 133.55,\ -12.42{]}}), and also the name of
variable nested in the \code{ts} column with its type:
\code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.

@@

\code{match_sites()} from \pkg{cubble} matches data based on these two
criteria. It first matches the two data sources spatially through
computing the pairwise distance based on latitude and longitude. Pairs
that pass the spatial matching are then matched temporally through
computing the number of matched peaks within a fixed length window.
Figure \ref{fig:illu-matching} illustrate this temporal matching in more
details. Given two series \code{A} and \code{a}. 3 peaks have been
picked in each series. An interval, with default length of 5, is
constructed for each peak in series \code{A} and the peaks in series
\code{a} are tested against whether they fall into the any of the
intervals. In this illustration, there are 2 matches for these two
series. Several arguments are available in \code{match_sites()} to
fine-tune the matching:

@@

Glyph map \citep{Wickham2012-yr} plots the time series as single glyph
on the map and in \proglang{R}, \pkg{GGally} implements the glyph map
through the \code{glyphs()} function. The function construct a data
frame with calculated position (\code{gx}. \code{gy}. \code{gid}. of
each point on the time series using linear algebra (Equation 1 and 2 in
\citet{Wickham2012-yr}). The data can then be piped into \code{ggplot}
to create the glyph map as:

@@

Polar glyph map can be specify as a parameter, \code{polar = TRUE}. in
the \code{geom_glyph()}. along with \code{width} and \code{height} in
either absolute or relative value. Global and local scale can be
controlled by the parameter \code{global_rescale}. which default to
\code{TRUE} for global scaling. Reference box and line can be added with
separate \code{geom_glyph_box()} and \code{geom_glyph_line()}.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # Assume my_ncdf has bounding box of [-180, -90, 180, -90] 
R> # at 0.25 degree resolution and subset it to have 
R> # 1 degree resolution:
R> dt <- as_cubble(my_ncdf, vars = c("q", "z"), 
+                 long_range = seq(-180, 180, 1), 
+                 lat_range = seq(-90, 90, 1))
\end{CodeInput}
\end{CodeChunk}

@@

With spatio-temporal data, users may wish to make plots to learn the
spatial distribution of a variable or to find patterns such as trend or
seasonality in the time series. Combining this two types of plot with
interactivity let users to link between points on the map and their
corresponding time series and explore the spatial and temporal dimension
of the data simultaneously. Below is an example that describes the
process of building an interactive graphic with \pkg{cubble} and
\pkg{crosstalk}.
